### 2018-11-13 paper/paper.rmd
start knitting paper

---
title: "Sending a Balloon to the Edge of Space"
author: Luker Bowsher, John Kim, Vivian Liu, Simon Oros, Caroline Pang, Alec Vercruysse
date: "November 13, 2018"
output: 
  pdf_document:
    number_sections: true
    keep_tex: true
    includes:
      in_header: "preamble.tex"
fontsize: "12pt"
indent: true
abstract: This paper details 6thsense, a mission to send a weather balloon to to edge of space in order to study Earth's atmosphere. The weather balloon is outfitted with a payload that includes a camera, inside and outside temperature sensors, a barometer, multiple GPS tracking devices, as well as a host of individual experiments. The balloon most likely reached an altitude of 20,000 meters, into the stratosphere and ozone layer. The balloon was successfully recovered after approximately one and a half hours of flight.
---

\newpage
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\tableofcontents
\newpage
\onehalfspacing

```{r setup, include=FALSE}
library(knitr)
library(mosaic)
library(gridExtra)
opts_chunk$set(echo = F)

```
#Introduction

There are four main layers to the atmosphere: the troposphere, stratosphere, mesosphere, and thermosphere. The **troposphere** is the section of the atmosphere from roughly 0km to 15km, and is the layer in which we live. It includes 75% of the mass of all gases in the atmosphere, consisting of 78% nitrogen gas, 21% oxygen, and less than 1% of argon, carbon dioxide, and other constituents. Because it is both the lowest in altitude and most massive layer, the pressure in the troposphere is the greatest: 100kPa. As you increase altitude in the troposphere, temperature decreases to about -60C. The **stratosphere** is the section from roughly 15km to 50km. The ozone layer, or the region in the atmosphere which absorbs most UV rays, is within the stratosphere at about 20km in altitude. The pressure in the stratosphere is 10kPa. The **mesosphere** is the region from 50km to 80km. The pressure in the mesosphere is 0.1 kPa. Finally, the **thermosphere** is the highest in altitude of the four layers, spanning from 80km into outer space. Here, the pressure is only 0.01kPa, as it is the least dense of the layers. The few air molecules in the thermosphere are heated to extremely high temperatures by the Sun’s radiation. However, because there are so few molecules, we cannot measure an appreciable difference in overall temperature, so the effective temperature of the thermosphere is less than that of the mesosphere.

The layers relevant to our experiment are the troposphere and stratosphere, as our balloon launch was designed to reach a maximum altitude of 30km (about the middle of the stratosphere but past the ozone layer). Therefore, we had to prepare the payload for temperatures as low as -60C and pressure as low as 10kPa (in the stratosphere). We expect to encounter the lowest temperature in the troposphere, around an altitude of 10km. As the balloon rises from sea level to the top of the troposphere, we expect to see temperature decrease because the ground absorbs most of the sun’s radiation (as opposed to the gas particles). However, as the balloon enters the stratosphere, we expect an increase in temperature due to the ozone layer’s blockage of UV rays. The UV rays that the ozone layer absorbs causes the gas particles in the stratosphere to heat up, so temperature increases with altitude. As the balloon gains altitude, the density of air around it decreases, so the pressure decreases.

The presence of greenhouse gasses in the atmosphere contribute to the warming of the earth’s surface and the troposphere because of the greenhouse effect. The greenhouse effect describes the trapping of heat within the earth’s atmosphere via the infrared absorption of certain molecules.

When infrared radiation enters the atmosphere, short-wave infrared radiation (700 -  1000 nm) does not effectively interact with greenhouse gasses. The photons instead pass through the atmosphere and reach the surface of the earth. Here, they are absorbed by the surface of the earth and get reemitted as a lower energy, longer-waved infrared photon. Greenhouse gasses can now absorb these lower energy photons and reemit them. This absorption and reemission warms the troposphere by increasing the kinetic energy the atmosphere. Additionally, greenhouse gasses can reemit photons back towards the surface of the earth, increasing the temperature of the earth’s surface. By impeding the tendency for long-wave radiation to leave the atmosphere, the greenhouse effect increase the temperature of the troposphere and surface.

The spectroscopy that explains the greenhouse effect concerns molecular vibration. Infrared photons excite vibrations in molecules. Since the vibrational frequency of atoms are quantized, only certain energies will excite vibration in a molecule. Different molecules can exhibit different vibrational modes. A molecule with more atoms has more vibrational modes. The infrared selection rule states that a vibrational mode is active when it is changing the dipole moment of the molecule. An active vibrational mode is a molecular vibration that can be excited by an infrared photon. Thus, infrared photons will only be absorbed when that molecule’s dipole moment is changing. Molecules such as nitrogen and oxygen will never exhibit a change in their dipole moment due to vibration because their only vibrational mode involves the singular oscillation of the distance between two atoms. Thus, they cannot absorb infrared radiation in order to partake in the greenhouse effect and are not considered greenhouse gasses. R On the other hand, many atmospheric molecules that have more than two atoms have more complex vibrational modes. These active vibrational mode gasses, such as nitrogen dioxide and water, are called greenhouse gasses.

As the greenhouse effect warms the planet, many factors accelerate the temperature rise by creating positive feedback loops. Positive feedback loops amplify any global increase in temperature. One positive feedback loops is created when rising temperatures melt glaciers. These white and reflective icy surfaces melt into dark oceans which absorb more radiation and heat up faster. This explains the approaching transition from a permanent to a seasonal ice cover in the Arctic Ocean. Thus temperature increases accelerate.^[https://www.nature.com/articles/s41598-017-08467-z] Another positive feedback loops involves the melting of permafrost. A quarter of the northern hemisphere is covered in permafrost that holds concentrated methane and carbon dioxide. Slight growths in temperature are enough to thaw the permafrost. When it melts, these greenhouse gasses are released into the atmosphere. Thus, the greenhouse effect amplifies and global temperature growth accelerates. ^[http://science.sciencemag.org/content/340/6129/183] Finally, humidity is also a positive feedback loop. As global temperatures rise, the atmosphere can hold more water vapor because less of it gets condensed. Water vapor is a greenhouse gas itself. Thus any increase in global temperatures will be amplified due to the water vapor positive feedback loop.

The annual increase of atmospheric carbon dioxide concentration is approximately 100 times greater in the past 60 years than it has been during any previous natural increase.F Additionally, methane concentrations have more than doubled since the start of the industrial revolution.B  Partly due to the increase of these gas concentration along with many positive feedback loops, the planet's average surface temperature has risen 0.9\degree C since the late 19th century.F With rising temperatures comes the consequences of severe weather patterns, flooding and other natural disasters. Collecting atmospheric data of things such as temperature and gas concentrations informs our understanding of the progression of the greenhouse effect. With this information, we can begin to look at ways to impede rising global temperatures.

The Panel on Climate Change has recognized that as global temperatures rise, lightning strikes will occur more frequently. As our skyscrapers get higher and multiply, lightning strike sourced Terrestrial Gamma ray Flashes (TGF) become a concern.

TGFs are short, powerful bursts of gamma rays detected in the atmosphere. Each burst contains 10^17 to 10^19 gamma rays and last about 1 ms. They were first detected in the 1950s. The United States, concerned with the Soviet Union’s nuclear progress used satellites to detect gamma rays and track the enemy’s progress. After observing long, 5-10 second gamma ray glows (suspected to be Gamma ray Bursts (GRB) from stars), the U.S. sent more satellites to explore GRB further. In addition to detecting GRB, the data from these satellites led to the discovery of TGFs. When short, gamma ray bursts were analyzed for their source locations, researchers noticed that these burst mainly came from coastal sites on the equator. After recognizing a correlation with lightning prone areas and these short gamma ray bursts, C. T. R. Wilson’s theory concerning a Relativistic Runaway Electron Avalanche (RREA) was applied to discover TGFs.

Wilson’s RREA theory relies on a quantum properties of electrons. As the velocity of normal objects traveling through the atmosphere increases, the air resistance increases also. Therefore, in a constant force field, the object will reach a terminal velocity. For electrons however, while the air resistance initially increase with its speed, at high velocities, the air resistance on the electron begins to decrease. This is because at relativistic speeds, the electron can avoid interactions with other molecules. Thus, after the electron reaches a certain speed, it will never reach a terminal velocity and accelerate indefinitely as long as it is still in a force field. Electrons that escape terminal velocity are called runaway electrons. The second part of Wilson theory involves an explanation of the “Avalanche.” Runaway electrons can knock other electrons while accelerating and cause them to become runaway electrons also. This creates an exponential increase of runaway electrons as more and more electrons interact.

The RREA theory explains a hypothesis for TGFs. The field that accelerates runaway electrons is an electric field produced by the separation of charges in thunderclouds. The energy field required to cause a RREA is powerful enough to cause lightning strikes. Thus, the occurrence of a lightning strike is correlated with that of a TGF. As electrons accelerate through the atmosphere, they can undergo Bremsstrahlung interactions with the nucleuses of air molecules. When an electron is deflected around a nucleus, a powerful Bremsstrahlung photon is emitted. Since an avalanche of electrons undergo this interaction, this explains the massive quantity of gamma rays contained in one TGF.

While there is an abundance of research done on upward directed TGFs (those detected by the satellites in the 1950s), there is a lack of data detecting downward directed TGFs. Currently, researchers led by David M. Smith at UC Santa Cruz and other researchers around the world are using a multitude of methods to detect downwards directed TGFs. Some methods include sending detectors in cargo planes, on weather balloons and on the ground in thunder prone areas. Since TGFs are so powerful, detectors usually fail if they are near a TGF because of the sheer amount of energy exerted. This field of research is relatively new and active.

**INSERT PRESSURE MODEL ANALYSIS HERE**

The exponential model for temperature vs altitude, where altitude is less than 11km (or where altitude is within the troposphere) is: T=T0(1-h/44329), where h=height in meters, and T0=temperature at sea level. This relationship shows a negative relationship between height and temperature. As altitude increases within the troposphere, temperature decreases.

# Experiments Conducted

The payload included two tracking devices, a GoPro Hero 3+ to record the ascent, and sensors to monitor the atmosphere and record altitude data. In addition, each team member implemented a sensor as part of an individual experiment. This paper focuses on the results of the group experiments, but the setups for the individual experiments are still described. Separate papers can be found for each team member's individual experiment.

## Group Experiments
```{r experiments_conducted, echo=FALSE}
group_experiments <- data.frame("Sensor" = c("Real Time Clock", "GPS", "Barometer", "Temperature", "Humidity", "Temperature"), "Function" = c("Accurate Timing", "Location Data", "Altitude Data", "Outside Temperature", "Outside Humidity", "Inside Temperature"), "Model" = c("Sparkfun DS1307 RTC", "Sparkfun Venus GPS", "Vernier Gas Pressure Sensor", "Adafruit BME280", "Adafruit BME280", "Thermistor"))
kable(group_experiments)
```

In addition, a PicoAPRS was used to provide live GPS coordinates, altitude, heading, and speed measurements.

These sensors were picked in order to be able to track the balloon through its flight and provide baseline measurements to compare other individual sensors to. By recording accurate altitude and time measurements with the measurement of every other sensor, sensor data can be analyzed with respect to both time and altitude. Furthermore, recording temperature data for both inside and outside the payload helps provide diagnostic data for if any specific part of the payload fails, and provides another baseline to compare any other sensor measurements to. The Humidity sensor not only provides humidity data which can be analyzed with respect to Altitude, but it can also determine when the payload is in a cloud, which could affect the measurements of other sensors.

## Individual Experiments Conducted
```{r inidivduals, echo=F}
individuals <- data.frame("Team Member" = c("Alec Vercruysse", "Luke Bowsher", "Simon Oros", "Vivian Liu", "Caroline Pang", "John Kim"), "Experiment" = c("Spectrometer", "Accelerometer", "~~Geiger Counter~~*", "Light Intensity Sensors", "UVB Sensor", "Methane Sensor"))
kable(individuals)
```

*shortly before launch it was discovered that the Geiger Counter was non-operable and therefore it was taken out of the payload. Instead, Simon Oros focused on ____.

**INCLUDE INDIVIDUAL MOTIVATIONS HERE**

# Physical Design

lorem ipsum

# Cut Down Mechanism

lorem ipsum

# Electrical and Software Design

lorem ipsum

# Tracking The Payload

One of the largest priorities was tracking the balloon to ensure a successful recovery. To do this, two completely independent tracking systems were implemented to add redundancy and mitigate the chances of completely losing the payload upon landing. The first method was a pre-built PicoAPRS transceiver, a small separately powered 144.39 MHz transceiver that utilizes an amateur radio network to broadcast location details that can be accessed through a website. The backup system was transmitting location data obtained by a Venus GPS module through a line-of-site, point to point, 900Mhz receiver, with a novel protocol developed with an emphasis on simplicity of parsing even broken packets. Lastly, to narrow down the location of the payload once it touched down, a separate transmitter at 147.065 MHz was included to use for fox-hunting upon touchdown.

## The Global Positioning System

The Global Positioning System (GPS) is a system of satellites in Medium Earth Orbit around Earth which provide accurate location details to any device connecting to the network. The orbits of these 27 satellites, while not geostationary, ensure that four satellites are visible from any location on Earth at any given moment.^[http://www.astronomy.ohio-state.edu/~pogge/Ast162/Unit5/gps.html#note01] By evaluating the time taken for GPS signals to be sent down to earth in the form of radio waves traveling at the speed of light, a GPS receiver is able to calculate the distance between it and the satellite. Using this distance information gained from at least three satellites, a GPS receiver is then able to “trilaterate” to find its location relative to the satellites.^[http://www.physics.org/article-questions.asp?id=55]
 Given the distance between a satellite, a GPS unit must be located on a sphere extending outward with a radius of that distance. With information from three satellites, three spheres are given, and generally the intersection of three spheres results in two points, once which can be discarded due to the fact that it would not be physically possible for the GPS receiver to be located there.

The Department of Commerce has placed regulations on exports that could potentially threaten the United States. This includes GPS receivers that can provide data when they are above 18,000 meters or their speed is above 1000 knots, due to their potential use in Intercontinental Ballistic Missiles.^[http://ravtrack.com/GPStracking/cocom-gps-tracking-limits/469/]Some manufacturers have implemented these limits by shutting down the system when either condition is met, and some manufacturers have implemented these limits by shutting down the system only if both conditions are met. This is an important distinction, since high altitude balloons often exceed the height limit of 18,000 meters. Sparkfun’s Venus GPS module, which uses the Venus634FLPx GPS chipset, is known to allow altitudes of over 18,000 meters, given that the speed of the GPS receiver is not calculated to be more than 1000 knots.^[https://ukhas.org.uk/guides:gps_modules]

## The PicoAPRS Transceiver

The primary tracking system onboard the balloon was a PicoAPRS, developed by Taner Schenker (DB1NTO). This is a small transceiver that can receive and broadcast on the Automatic Packet Reporting System (APRS). Most importantly, it contains its own separate GPS chip that allows it to broadcast its location to the APRS network. This device is completely self contained, including its own 850MAh Lithium Ion battery, making it a perfect candidate for tracking -- even if all other systems on the balloon have failed, the PicoAPRS should still be able to transmit its location. Furthermore, the PicoAPRS provides a simple menu for device configuration, to enable the performance desired. Specifically, the PicoAPRS transmitted GPS beacons at a power of 1 Watt over intervals of 60 seconds. The PicoAPRS did this with Menlo School’s callsign, N6MLO, and a unique SSID, 9, that differentiated the packets sent by 6thsense’s PicoAPRS from other concurrent Menlo School balloon launches.

## The APRS Network and MIC-Encoding

A very important benefit of using the PicoAPRS is that it uses the Automatic Packet Reporting System (APRS). This amateur radio network, developed by Bob Bruninga (WB4APR) uses digipeaters to forward properly formatted broadcasted beacon packets to other stations in the area to effectively extend the range of a single transmitter. Rather than focusing on ensuring that a packet is received by all digipeaters, this protocol focuses on redundancy through packet “multiplication” by having each digipeater that receives the signal propagate it outward through the network. While the exact details of how many “hops” a packet can take through the network before transmission ends are dependent on the settings present in the header of the packet. This system became popular, especially for transmitting GPS data and tracking vehicles, and eventually specifications were developed to transfer packet data to the web through APRS-Internet Service (APRS-IS).^[http://www.aprs-is.net/] The aprs.fi website, which interfaces with APRS-IS and displays live telemetry data superimposed on google maps, provided a simple method of tracking the balloon during the launch.

The APRS network transmits packets through a single AX.25 data link protocol, which specifies the structure of frames to be sent through the network. Specifically, it uses Unnumbered Information (UI) frames, a type of AX.25 frame that is transmitted without any expectation of a response confirming reception, and reception is not guaranteed. APRS supports different encoding of these AX.25 UI-frames, however, to allow for different uses of the network. The PicoAPRS transmitter uses “Mic-Encoding” (MIC-E), a method that allows for extremely compressed packets that still contain position, course, speed, a message, and all relevant path settings and necessary headers. This achieved by writing compressed latitude information in the destination address field of the AX.25 frame, and compressed Longitude information in the frame’s Information Field. By transmitting extremely short packets that are still supported by the APRS network and aprs.fi, Mic-E can improve the reliability of the beacons transmitted.

APRS is transmitted using Frequency Modulated signal at 144.39 MHz. Frequency Modulation works by altering the frequency of a carrier wave to encode information.

The PicoAPRS was chosen because the system had worked historically for past ASR Balloon Launches. During testing, however, the APRS network often failed to receive beacons, due to the fact that most tests were conducted at ground level, where line of sight to repeater stations was often not available. Furthermore, there was no method of testing the maximum altitude that beacons could be sent from before connection to the APRS network was lost. Lastly, since Lithium-Ion batteries are known to work poorly in below freezing temperatures, there were concerns that the unit might shut down once the temperature dropped, if not properly insulated. Due to these concerns, a backup tracking system was developed that did not rely on the PicoAPRS and APRS network should the PicoAPRS unit fail.

## The 900MHz Venus GPS Transciever

The backup tracking system involved sending GPS along with other telemetry data from the main arduino mega to the base station using Digikey 9XTend 900MHz transceivers.^[https://www.sparkfun.com/datasheets/Wireless/Zigbee/xtend-productmanual.pdf] The use of these transceivers provided a high power 1 Watt signal in a relatively clear band that let a novel communications protocol be implemented that allowed for ease of parsing. The 9XTend module interfaced easily with both balloon and base station arduinos through Universal Asynchronous Receiver-Transmitter (UART) Serial. Since the Arduino Mega provides three usable Serial ports, one was used to send data to the 9XTend Module. By taking advantage of Arduino’s Serial library, a library was developed that allowed the 9XTend modules to transmit ASCII encoded bytes, which, given some stream editing by the base station to remove ASCII control characters injected by the base 9XTend Module, is parsable as an ASCII byte stream by software on the base station.

### A Novel Transmission Protocol

To send readable GPS and telemetry data to the base station, a simple multi-layer protocol was implemented with design considerations in mind to allow for simple parsing and data analysis, simple manual error detection, no error correction or discarding of broken transmissions, no set packet size, zero unnecessary headers or other packet configuration information, and no transmitter handshaking requirements. This allows for simplex communication between the payload and base station. This also ensures that all bytes that are received by the 9XTend, which implements a Cyclic Redundancy Check on every byte received to ensure its value is correct, are sent to the base station software for parsing. Note that the 9XTend module not protect against lost bytes, so packets could often be broken or missing some information. Through this new protocol, no partially broken packets would be lost or deleted if the signal was weak. Since the 9XTend Module transmitted all data for sensors connected to the arduino as a redundancy in case the payload was not retrieved or there was a problem logging data to the onboard SD card, it is also crucial to be able to differentiate between two different values in a partially broken packet.

A protocol was used with the following form to transmit n values:

`******<value-1>------<value-2>------...------<value-n>`

Where `<value-i>` is an ASCII encoded, calibrated and human readable sensor value. This allows for simplex operation in which a base station can tune in at any time--if the base station loses reception or power for some reason, it can listen in for the next packet when it comes back online by simply waiting for the 6 stars indicating the start of a packet. Furthermore, data was encoded in ASCII bytes to let the output be human readable with little computation. This way, the output stream could be easily analyzed without having to convert bytes containing raw numerical values into ASCII text. This gets rid of the need to make specifications for negative numbers and floats, and the need to specify the size of transmitted data. By encoding each digit as an ASCII byte, it was simple to account for negative numbers by preceding the value with a dash just as is normally done with displayed numbers, and to account for floats by adding a decimal place behind the ones digit. Furthermore, more complicated sensor readings like GPS and Spectrometer output that contained ASCII formatted text could be sent as-is with zero need for manipulation. Since 12 values were transmitted (refer to Section 5) with uncompressed ASCII bytes, packet sizes were long: up to 240 bytes. Because of this, there were often small failures in the packet were a byte or series of bytes was not received. Such packets were still logged with the notion that it was still possible to extract most of the data out of them instead of having them completely go to waste. To ensure that the start of a packet could be recognized and separate values could be delineated properly, however, a full six bytes was used to mark the beginning of a transmission, and breaks separating values. This data was only transmitted about once every 12 seconds, partly to serve as flow control in order to not overload the transmitter’s buffer, and partly to conserve battery power, as the 9XTend module consumed 1 Watt while transmitting and was connected to the same battery as the Arduino Mega running all the sensors and SD logger.

### Physical Considerations and Antenna Design

Since the 9XTend module transmits directly to the base station, line of sight was crucial to maintain connectivity. An omnidirectional whip antenna had to be used on the balloon module since the heading and position relative to the base station was constantly changing and therefore the transmissions could not be directed in one particular direction. While it might have been possible to use a directional antenna at the base station to increase signal, it was not ideal since the balloon was followed in a chase van, and directing an antenna outside the van would be imprecise and lead to a lossy connection. Furthermore, 900MHz yagi antennas are not readily available in stores and time was not properly budgeted to assemble and test one by hand. Therefore, a 900Mhz whip antenna was mounted to the top of the chase van to connect to the receiving 9XTend Module.

## Fox-Hunting with a 147.065MHz Transmitter

While GPS modules, if they have signal at ground level, are usually able to report location down to a few feet, the PicoAPRS was only expected to be able to connect to the APRS network at a few thousand feet before it lost line of sight to repeater stations, and there was no expectation to receive signal from the 900MHz point to point transmitter, which also relied on line of sight, which would not be available when the payload had touched down. Therefore, it was expected that the latitude and longitude data provided by both the PicoAPRS and the 9XTend modules would be too imprecise to be able to locate and retrieve the payload. To further narrow down the location of the module, a 147.065MHz transmitter was included that broadcasted a distinct audio beacon along with Menlo School’s callsign in morse code, every minute. This enabled the use of Fox-Hunting techniques to recover the payload once the tracking team was in range of the beacon.

By having an omnidirectional whip antenna transmit the beacon and a directional yagi antenna receive the beacon, the direction of the balloon relative to the fox-hunt receiver was able to be calculated once the fox-hunt receiver was in range of the beacon. Since directional antennas receive greater power in the direction they’re pointed, by identifying in which direction the antenna receives the strongest signal, it is possible to identify the source of the beacon and move in that direction until the payload is found.

### Fox-Hunting Yagi Antenna Design

**CAROLINE STILL NEEDS TO WRITE**

### Determining Signal Strength with a Variable RF Attenuator

A crucial component of Fox-Hunting is measuring the received signal strength. Since the transmitter transmits an audio signal, one can simply listen to the clarity of the received audio to do so. If the receiver is close to the transmitter, however, and the received signal is very strong, in tests it was hard to differentiate between a weak and a strong signal and narrow down on the proper direction. To be able to better distinguish signal strength, a variable Radio Frequency (RF) attenuator was used to lower the strength of the signal until distortion in the audio could clearly be heard, and therefore it was easier to determine whether the signal strength improved or decreased after a change in direction of the antenna.

The RF Attenuator used to lower the received signal mixes a constant 4MHz with the received signal in order generate two sidebands 4MHz above and below the base signal. This sideband is an attenuated version of the original signal, and tuning the receiving radio to the base frequency plus 4 MHz, 151.065MHz, lets one monitor this attenuated signal. Furthermore, the amplitude of the 4 Mhz signal can be varied in order to vary the strength of the sideband, effectively providing varied attenuation.

## Results

The balloon powered on at precisely 1:00pm with all three systems working: location information showed up on aprs.fi, the base station received transmissions including GPS information, and the HAM radio used for fox-hunting received 6thsense’s unique tone.

### The Venus GPS and 9XTend Module

Only about 14 minutes later, however, once the balloon had reached approximately 2600 meters, the Venus GPS module lost reception and stopped transmitting latitude and longitude data. The 9XTend module continued to transmit other telemetry data including altitude, however, for the duration of the flight, until the payload touched down, in which most likely the crash disconnected some circuitry essential to transmission, or the batteries lost their charge. Once the payload reached approximately 7800 meters, 9XTend transmissions became significantly damaged to the point where the data could no longer be programmatically parsed by simply finding values between the delimiters specified. Packets continued to be somewhat recognizable up to 20,000 meters, the estimated maximum height the ballon reached at approximately 1:50pm. Shortly after, at 1:53pm, the first packet was received indicating the balloon had began a slow descent (see Section 4). During descent, especially back at around 7800 meters, transmissions became more recognizable but still required manual interpretation. Transmission stopped after around 7400 meters, indicating that most likely the arduino lost power at 2:17pm, two and a quarter hours after powering up.

Unfortunately, while the 9Xtend module paired with the Venus GPS failed its primary mission of providing GPS information, the transmission system worked very well for the entirety of the time that it was powered. Not only did it provide the base station with live altitude data, it transmitted sensor data, including that of individual experiments. This proved invaluable to the team, since the SD logging module failed to write data.

Since the Venus GPS receiver was mounted at the top of the payload, with its line of sight to the GPS satellites obstructed by only the balloon, and the GPS chips did not have the 18,000 meter limit, there is no clear reason for why the system failed. One possibility is that the outside temperature of 16 degrees Celsius was too low for the circuitry involved. Another is simply that the antenna was somehow internally damaged or not optimal. Upon touchdown, all GPS wiring was still in place, indicating that it was not an issue with the circuit or placement of the antenna.

The last possible reason is that other circuitry such as the arduino microprocessor in the payload generated radio frequency interference. The GPS center frequency is 1575.42MHz, however, and the crystal oscillator providing the clock for the Arduino Mega only runs at 16MHz, which indicates that other circuits in the payload were most likely not causing this issue.

### The PicoAPRS and APRS Network

The PicoAPRS consistently broadcasted packets that were picked up by the network up to almost exactly 10,000 meters at 1:30pm, before the network lost reception. The PicoAPRS continued transmitting accurate location data, however, and the network once again started receiving transmissions at 2:16pm, once the payload had fallen to approximately 8200 meters. The final transmission came at 2:36pm just 250 meters above the ground. Overall, this system worked flawlessly to provide accurate location data while in range of the network, between approximately 250 meters and 9000 meters in altitude. Fortunately, the insulation added to the payload was able to keep its Lithium Ion battery warm enough to maintain power. Since the Venus GPS failed early into the flight, tracking the payload relied on the APRS for accurate location data until the chase van got in range of the fox-hunt beacon .

### The Fox-Hunt Beacon and Yagi Antenna

The final PicoAPRS beacon came from a very low altitude, narrowing down the final location to a radius of less than 200 meters. Once the chase van arrived at the location described by the final APRS packet, the team was immediately able to pick up the Fox-Hunt signal even with the attenuator attached. A single sweep of the Yagi antenna revealed the payload was due South along the road, and the payload was found less than 50 meters away from the road, lying hidden from view of the road on a small hill. The antenna, attenuator, transmitter and receiver all worked flawlessly to lead to a simple recovery in the time it took to walk from the parked van to the touchdown point of the payload.

# Sensor Calibration

To ensure that the payload would collect accurate and reliable data, the lightweight sensors used and controlled by the Arduino board were thoroughly calibrated. The three group sensors calibrated were the temperature and humidity sensors onboard the Vernier Gas Pressure Sensor, and the Vernier Gas Pressure Sensor. The outputs of all three sensors were compared to standards, which were already known to be reliable and accurate. The standards, despite their accuracy, could not be used in the payload because of their heavy weight and large size, hence it was crucial that the small and light Vernier and BME280 sensors were calibrated.

## The Vernier Gas Pressure Sensor

The Vernier Gas Pressure Sensor, model GPS-BTA, was calibrated to the Extech SD700 Barometric Pressure/Humidity/Temperature Datalogger, which was used as our standard for pressure. The Vernier sensor was connected to an Arduino microcontroller with a Vernier Analog Protoboard Adaptor and configured to log data onto a micro-SD card every 1.50 seconds. Instead of having a output unit of hPa or kPa, the Arduino maps input voltages between 0 to 5 volts to 10-bit integer values between 0 and 1023—a total of 1024 bins with 0.49 mV resolution. Thus, each unit represents 4.9 mV. The SD700 has a pressure range of 10 to 1100 hPa, fine resolution of 0.1 hPa, optimal temperature range of 0 to 50°C, response time of 10 milliseconds, and total accuracy with its factory calibration ±4 kPa. It requires 6 AAA batteries, and records data to an external SD card in the format of an Excel worksheet. It was configured so that it sampled data every five seconds to have smoother and more gradual changes in pressure. After both sensors were turned on at the same time to easily compare the data later, they were immediately placed in a vacuum chamber, in which air and other gasses are removed by a vacuum pump, thus creating a low-pressure environment inside the chamber. The release valve was closed and the vacuum pump was opened. When the pressure reading from the valve reached its minimum of approximately 0 hPa, the vacuum pump was then disconnected. An important part of the pressure calibration process was having “plateaus,” or multiple data points at various pressure levels. To have plateaus in the data, the release valve was slowly opened and closed right after, and this process was repeated after waiting approximately 45 seconds multiple times. Graph 5 and 6 show the varying pressure readings from the standard Extech sensor and Arduino Vernier sensor across time, respectively.

```{r vernier_calibration, echo=F}
th <- theme(plot.title = element_text(size=10))

pressure1 <- read.csv("data/pressure1.csv")
p1 <- ggplot(pressure1, aes(x=pressure1$Time, y=pressure1$Arduino.output)) +th + geom_point() + labs(title="Graph 5: Varying pressure readings from \nstandard (Extech) sensor across time", x="Time (s)", y="Pressure (hPa)")

pressure2 <- read.csv("data/pressure2.csv")
p2 <- ggplot(pressure2, aes(x=pressure2$Time, y=pressure2$hPa)) + geom_point() + th + labs(title="Graph 6: Varying pressure readings from \nArduino (Vernier) sensor across time", x="Time (s)", y="Pressure (10-bit unit)")

grid.arrange(p1, p2, ncol=2)
```

Graph 7, shown below, shows the conversion between the output of the standard (Extech) sensor and that of the Arduino (Vernier) sensor, including the equation to convert the Arduino’s 0.49 mV resolution units to kPa.

```{r vernier_cal2, fig.height=3.5}
pressure3 <- read.csv("data/pressure3.csv")
names(pressure3)[1] <- "Vernier sensor (10-bit unit)"
names(pressure3)[2] <- "Extech sensor (kPa)"

lm_eqn_p3 <- function(df){
    m <- lm(df$`Extech sensor (kPa)` ~ df$`Vernier sensor (10-bit unit)`, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3), 
              b = format(coef(m)[2], digits = 3), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}

p3 <- ggplot(pressure3, aes(x=pressure3$`Vernier sensor (10-bit unit)`, y=pressure3$`Extech sensor (kPa)`)) + geom_point() + labs(title="Graph 7: Pressure conversion between\n standard (Extech) and Arduino (Vernier) sensor", x="Pressure from Vernier (10-bit unit)", y="Pressure from Extech (kPa)") + geom_smooth(method = "lm") + geom_text(x=300, y=90, label=lm_eqn_p3(pressure3), parse=TRUE)
p3

```

`r kable(pressure3, caption="Varying average pressure readings from Arduino (Vernier) vs. Standard (Extech) sensor")`

## The Adafruit BME280 Sensor

The Adafruit BME280 sensor, which takes data on temperature, barometric pressure, and humidity, was calibrated to a Vernier Relative Humidity Sensor for humidity and the Fieldpiece ST4 Dual Temperature Meter for temperature. To calibrate for humidity, the BME280 was connected to an Arduino mega with header pins and programmed to print out data every second to the serial monitor. The standard for humidity, a Vernier humidity probe,  was connected to the same Arduino board via the Analog Protoboard Adaptor just like the Vernier Gas Pressure Sensor and also programmed to print out its humidity readings in the serial monitor every second. The Vernier Relative Humidity Sensor has a humidity range of 0% to 95%, resolution of 0.16% RH, operating temperature range of 0 to 85°C, response time of 40 seconds, and total accuracy with its standard calibration of ±10% RH. Both sensors were exposed to environments with various humidity levels: the Menlo Quad at 3:00 PM and a shower 30, 60, 90, and 120 seconds after turning the hot water on. Multiple consecutive relative humidity readings from both sensors at each time point were then manually recorded in a csv file. Graph 8, shown below, depicts the relative humidity conversion between the output of the standard (Vernier) sensor and that of the Arduino (BME280) sensor across various humidity levels.

```{r bme, fig.height=4}

humidity1 <- read.csv("data/humidity1.csv")
lm_eqn_h1 <- function(df){
    m <- lm(df$Arduino.... ~ df$Standard...., df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3), 
              b = format(coef(m)[2], digits = 3), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}
h1 <- ggplot(humidity1, aes(x=humidity1$Arduino...., y=humidity1$Standard....)) + geom_point() + labs(title="Graph 8: Humidity conversion between\n standard (Vernier) and Arduino (BME280) sensor", x="Relative humidity from BME280 (% RH)", y="Relative humidity from Vernier (% RH)") + geom_smooth(method = "lm") + geom_text(x=55, y=60, label=lm_eqn_h1(humidity1), parse=TRUE)
h1
```

In order to calibrate the temperature readings of the BME280 to a standard, the Fieldpiece ST4 Dual Temperature Meter, both sensors were placed in multiple locations of varying temperatures: a freezer, refrigerator, and the Menlo Quad at two different times of the day.

```{r bme_temp, fig.height=4}
temp1 <- read.csv("data/temp1.csv")
lm_eqn_t1 <- function(df){
    m <- lm(df$Arduino...C. ~ df$Fieldpiece...C., df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 3), 
              b = format(coef(m)[2], digits = 3), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}
t1 <- ggplot(temp1, aes(x=temp1$Arduino...C., y=temp1$Fieldpiece...C.)) + geom_point() + labs(title="Graph 9: Temperature conversion between\n standard (Fieldpiece) and Arduino (BME280) sensor", x="Temperature from BME280 (°C)", y="Temperature from Fieldpiece's T1 rod (°C)") + geom_smooth(method = "lm") + geom_text(x=5, y=20, label=lm_eqn_t1(temp1), parse=TRUE)
t1
errors <- read.csv("data/errors.csv")
names(errors)[2] <- "Percent error (%)"
```

`r kable(errors, caption="Percent errors for the group sensors used relative to standards")`

# Experimental Results

lorem ipsum

# Conclusion

lorem ipsum