### 2018-11-14 paper/bibliography.Rmd
almost done with paper

---
title: "Bibliography"
output: pdf_document
header-includes:
  - \AtBeginDocument{\let\maketitle\relax}
  - \usepackage{setspace}
indent: false
---
#References

*9XTend OEM RF module* [Product Manual]. (n.d.). Retrieved November 14, 2018, from

|     https://www.sparkfun.com/datasheets/Wireless/Zigbee/xtend-productmanual.pdf

&nbsp;

Bruninga, B. (n.d.). Automatic packet reporting system. Retrieved November 14, 2018, from

|        http://aprs.org

&nbsp;

Bruninga, B. (n.d.). Kits: APRS MIC-Encoder. Retrieved November 14, 2018, from

|        https://www.tapr.org/kits_mic-e.html#project

&nbsp;

Buthod, B. (Ed.). (1997, December 27). AX.25 amateur packet-radio link-layer protocol. Retrieved November 14, 2018, from

|        https://www.tapr.org/pub_ax25.html

&nbsp;

*COCOM GPS tracking limits* [Fact sheet]. (n.d.). Retrieved November 14, 2018, from

|        http://ravtrack.com/GPStracking/cocom-gps-tracking-limits/469/

&nbsp;

*GPS Modules* [Fact sheet]. (2016, September 5). Retrieved November 14, 2018, from

|        https://ukhas.org.uk/guides:gps_modules

&nbsp;

GPS Systems: Space segment. (2018, September). In GPS.gov. Retrieved November 14, 2018, from

|        https://www.gps.gov/systems/gps/space/

&nbsp;

Greenhouse effect. (2018). In *Encyclopædia Britannica*.

&nbsp;

How does GPS work? (n.d.). Retrieved November 14, 2018, from

|        http://www.physics.org/article-questions.asp?id=55

&nbsp;

Kashiwase, H. (2017, August). *Evidence for ice-ocean albedo feedback in the Arctic Ocean shifting to a seasonal ice zone*. Retrieved from Nature website:

|        https://www.nature.com/articles/s41598-017-08467-z

&nbsp;

Lindsey, R. (2018, August 1). Climate Change: Atmospheric Carbon Dioxide. Retrieved November 14, 2018, from

|        https://www.climate.gov/news-features/understanding-climate/climate-change-atmospheric-carbon-dioxide

&nbsp;

Loveall, P. (n.d.). APRS-IS [Fact sheet]. Retrieved November 14, 2018, from http://www.aprs-is.net/

&nbsp;

Pecsok, R. L., & Shields, L. D. (1968). *Modern Methods of Chemical Analysis*. New York: Wiley.

&nbsp;

Smith, D. M. (2015). *Terrestrial Gamma-Ray Flashes, Relativistic Runaway, and High-Energy Radiation in the Atmosphere*. Santa Cruz, CA: UC Santa Cruz.

&nbsp;

UCAR. (2011). *The Greenhouse Effect*. Retrieved November 14, 2018, from UCAR website:

|        https://scied.ucar.edu/longcontent/greenhouse-effect

&nbsp;

UCAR. (2012). Molecules Vibrate. Retrieved November 14, 2018, from UCAR website:

|        https://scied.ucar.edu/molecular-vibration-modes

&nbsp;

Vaks, A. (2013, April). *Speleothems Reveal 500,000-Year History of Siberian Permafrost.* Retrieved from Science website:
|        http://science.sciencemag.org/content/340/6129/183

&nbsp;

Wade, I. (Ed.). (2000, August 29). *APRS protocol reference* [White paper]. Retrieved November 14, 2018, from

|        http://www.aprs.org/doc/APRS101.PDF

&nbsp;

Yagi antenna theory: Yagi antenna basics. (n.d.). Retrieved November 14, 2018, from

|        https://www.electronics-notes.com/articles/antennas-propagation/yagi-uda-antenna-aerial/theory.php

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\setlength{\parskip}{8pt}