### 2018-11-14 paper/final_comp.Rmd

---
title: "final_comp"
output: pdf_document
header-includes:
  - \AtBeginDocument{\let\maketitle\relax}
  - \usepackage{pdfpages}

---
\includepdf[pages={-}]{paper.pdf}
\includepdf[pages={-}]{bibliography.pdf}
\includepdf[pages={-}]{bibliography.pdf}

# Appendix II: Table Comparing Atmospheric Composition vs Altitude

```{r appendix2, echo=F}
knitr::include_graphics("assets/appendix2.png")
```

\newpage

# Appendix III: Layers of the Atmosphere
```{r appendix3, echo=F}
knitr::include_graphics("assets/appendix3.png")
```

\includepdf[pages={-}]{assets/cad1.pdf}
\includepdf[pages={-}]{assets/cad2.pdf}
\includepdf[pages={-}]{assets/cad3.pdf}
\includepdf[pages={-}]{assets/cad4.pdf}